Speed and strength of an epidemic intervention

### 2020-10-09 speedstrength.tex
post-meeting edits

\begin{figure}[!t]
\includegraphics[width=\textwidth]{code/condom.Rout.pdf}
\caption{
\textbf{Evaluating a condom intervention using strength-based decomposition.}
(A) Condom use is thought to reduce probability of transmission by a similar factor throughout the course of infection; thus the multiplicative reduction $\tsub{L}{condom}$ due to condom use is constant across the course of infection.
(B) The estimated amount of early transmission affects estimated strength of the epidemic but not of a condom-based intervention.
The black and red circles indicate the baseline scenario.
}
\figlab{condom}
\end{figure}

%% \jd{For next version, consider making \figref{condom} a four-panel figure like \figref{test}.}

\begin{figure}[!t]
\includegraphics[width=\textwidth]{code/tt.Rout.pdf}
\caption{
\textbf{Evaluating a test-and-treat intervention using strength- and speed-based decomposition.}
(A) The strength of the test-and-treat intervention (calculated from the assumed hazard, (C)). The dashed line shows the corresponding effective strength of the intervention (from \eqref{strengthMean}) assuming 23\% early transmission.
(B) Increase in the estimated amount of early transmission decreases the estimated strength of an epidemic as well as the estimated strength of test-and-treat intervention.
(C) The assumed hazard for the test-and-treat intervention. 
The dashed line shows the corresponding effective speed of the intervention (from \eqref{speedMean}) assuming 23\% early transmission.
(D) The estimated amount of early transmission has little effect on the effective speed of intervention, and none on the speed of the epidemic estimated from incidence data.
Circles indicate the baseline scenario.
Test-and-treat intervention is modeled phenomenologically: $\tsub{L}{test}(\tau) = \exp\left(\int_0^\tau \tsub{h}{test}(\sigma) d\sigma \right)$ and $\tsub{h}{test}(\tau) = \tsub{h}{max} (1 - \exp(- K f(\tau)))$, where $f(\tau)$ is a gamma probability density function with a mean of $1$ year and a shape parameter of 2, $K = 4/\max(f(\tau))$, and $\tsub{h}{max} = 2\,\mathrm{year}^{-1}$.
}
\figlab{test}
\end{figure}

@@

For interventions which focus on infected individuals, like contact tracing or test-based isolation, the speed-based paradigm provides clearer insight into the likely effects of early transmission. 
More early transmission does not change our estimate of the speed of the epidemic, but it does change our estimate of the \emph{effective} speed of a given control curve.

Early transmission gives more weight in the calculation of effective speed ($\phi$) to the early period of infection.
Thus, if the hazard of control is increasing through time, as might be expected with isolation of symptomatic cases, more early transmission means that the effective speed of control is lower than originally thought, and control will be harder.
Conversely, if hazard of control decreases through time, as might be expected with isolation of returning travellers, or event-based contact tracing, more early transmission means that control will be easier than originally though.

With sufficiently detailed assumptions, we could do a correct calculation for any scenario in any paradigm. 
But in these examples, using the more appropriate paradigm for each scenario lets us know what to expect, and may strengthen our intuition for \emph{how} the assumptions lead to the result.

@@

In this study, we:
first extended the standard paradigm of \RR\ as critical parameter for control, by defining the strength of an intervention on the same scale as \RR, the strength of the epidemic; 
then constructed a parallel interpretation which measures the speed of an intervention on the same scale as $r$, the speed of an epidemic.
We thus showed that the standard paradigm for \RR\ and control has a natural parallel interpretation in terms of $r$.

@@

@@

When comparing interventions with epidemic parameters to evaluate strategies, the situation is similar. Some scenarios lend themselves naturally to a single approach. For example, in the classic case of vaccination to eliminate a previously established childhood disease, both disease spread and intervention can be clearly characterized using strength \citep{anderson1985vaccination}. In our HIV example, both the HIV epidemic and the test-and-treat intervention can be best characterized using speed. Other cases, such as using social distancing (a strength-like intervention) in the early stages of COVID-19 (epidemic speed is observed) may not fit so neatly into either paradigm, however.

>>>>

\begin{figure}[!t]
\includegraphics[width=\textwidth]{code/condom.Rout.pdf}
\caption{
\textbf{Evaluating a condom intervention using strength-based decomposition.}
(A) Condom use is thought to reduce probability of transmission by a similar factor throughout the course of infection; thus the multiplicative reduction $\tsub{L}{condom}$ due to condom use is constant across the course of infection.
(B) The estimated amount of early transmission affects estimated epidemic strength \RR, but not intervention strength \Rhat of a condom-based intervention.
The black and red circles indicate the baseline scenario.
}
\figlab{condom}
\end{figure}

\jd{For next version, consider making \figref{condom} a four-panel figure like \figref{test}.}
\jd{The above is an old comment. I'm not currently loving the idea of including that figure, but I guess we should make it and see?}

\begin{figure}[!t]
\includegraphics[width=\textwidth]{code/tt.Rout.pdf}
\caption{
\textbf{Evaluating a test-and-treat intervention using strength- and speed-based decomposition.}
(A) The strength of the test-and-treat intervention (calculated from the assumed hazard, (C)). The dashed line shows the corresponding effective strength of the intervention (from \eqref{strengthMean}) assuming 23\% early transmission.
(B) Increase in the estimated amount of early transmission decreases the estimated epidemic strength \RR\ as well as the estimated strength of a test-and-treat intervention \Rhat.
(C) The assumed hazard for the test-and-treat intervention. 
The dashed line shows the corresponding effective speed of the intervention (from \eqref{speedMean}) assuming 23\% early transmission.
(D) The estimated amount of early transmission has little effect on the effective speed of intervention \rhat, and none on the speed of the epidemic estimated from incidence data \rr.
Circles indicate the baseline scenario.
Test-and-treat intervention is modeled phenomenologically: $\tsub{L}{test}(\tau) = \exp\left(\int_0^\tau \tsub{h}{test}(\sigma) d\sigma \right)$ and $\tsub{h}{test}(\tau) = \tsub{h}{max} (1 - \exp(- K f(\tau)))$, where $f(\tau)$ is a gamma probability density function with a mean of $1$ year and a shape parameter of 2, $K = 4/\max(f(\tau))$, and $\tsub{h}{max} = 2\,\mathrm{year}^{-1}$.
}
\figlab{test}
\end{figure}

@@

For interventions which focus on infected individuals, like contact tracing or test-based isolation, the speed-based paradigm provides clearer insight into the likely effects of early transmission. 
More early transmission does not change our estimate of the speed of the epidemic, but it does change our estimate of the \emph{effective} speed of a given control curve. 
This is because more early transmission gives more weight in the calculation of effective speed ($\phi$) to the early period of infection.
Thus, if the hazard of control is increasing through time, as might be expected with isolation of symptomatic cases, more early transmission means that the effective speed of control is lower than originally thought, and control will be harder.
Conversely, if hazard of control decreases through time, as might be expected with isolation of returning travellers, or event-based contact tracing, more early transmission means that control will be easier than originally though.

@@

In this study, we:
first extended the standard paradigm of \RR\ as a critical parameter for control, by defining the strength of an intervention on the same scale as \RR, the strength of the epidemic; 
then constructed a parallel interpretation which measures the speed of an intervention on the same scale as $r$, the speed of an epidemic.
We thus showed that the standard paradigm for \RR\ and control has a natural parallel interpretation in terms of $r$.

@@

We also discussed the question of uncertainties introduced by the unknown magnitude of pre-symptomatic transmission of COVID-19. We showed that the strength-based paradigm provides clear insight into how this uncertainty affects interventions targeted to the general population, while speed-based paradigm is a clearer way to think about interventions targeted to infected people.

@@

When comparing interventions with epidemic parameters to evaluate strategies, the situation is similar. Some scenarios lend themselves naturally to a single approach. For example, in the classic case of vaccination to eliminate a previously established childhood disease, both disease spread and intervention can be clearly characterized using strength \citep{anderson1985vaccination}. In our HIV example, both the HIV epidemic and the test-and-treat intervention can be best characterized using speed. Other cases, such as using social distancing (a strength-like intervention) in the early stages of COVID-19 (epidemic speed is observed) may not fit so neatly into either paradigm, however.
With sufficiently detailed assumptions, we could do a correct calculation for any scenario in any paradigm. 
But in many of these examples, using the more appropriate paradigm for each scenario lets us know what to expect, and may strengthen our intuition for \emph{how} the assumptions lead to the result.
